---
title: "Reporting practices of statistical results"
output: html_document
date: "2025-08-11"
---

### Load packages
```{r include=FALSE}
library(metacheck) 
library(dplyr) 
library(readxl)
library(here)
library(rlang)
```

Steps 1, 2, and 3 do not need to be redone. You can proceed directly to Step 4 to load the RDS object containing the articles.

### Step 1- Convert all PDFs to XML format and save
```{r}
# List files in "pdf_files"
pdf_files <- list.files(
  path = here("files", "pdf_files"),
  pattern = "\\.pdf$", 
  full.names = TRUE
  )

# Convert PDFs to XML format
pdf2grobid(
  pdf_files, 
  save_path = here("files", "xml_files"), 
  grobid_url = "https://thesanogoeffect-grobid-papercheck.hf.space"
)
```
Two PDFs could not be converted to XML format resulting in a total of 347 studies in XML format.

### Step 2- Read in XML files
```{r}
# List files in "xml_files"
path_files <- list.files(
  path = here("files", "xml_files"), 
  pattern = "\\.xml$", 
  full.names = TRUE
)

# Read in XML files
articles <- read(path_files)
```

### Step 3- Store XML files as an RDS object for efficiency
```{r}
saveRDS(articles, "articles_xml.Rds")
```

### Step 4- Load RDS object
```{r}
articles_files <- readRDS("articles_xml.Rds")
```

## Run metacheck modules
### `all_p_values` module 

It returns all reported p-values in the articles. 
```{r}
all_p <- module_run(
  articles_files, 
  "all_p_values"
  )

all_ps <- all_p$table |>
  count(text, sort = TRUE)

all_ps
```


### `stat_p_exact` module

It returns exact significant p-values
```{r}
exact <- module_run(
  articles_files, 
  "stat_p_exact"
  )

exact$table |> 
  count(text, sort = TRUE)

tbl <- exact$table |> 
  filter(section == "results") |> 
  nrow()
```

### `stat_p_nonsig` module

It returns exact non-significant p-values
```{r}
exact_nonsig <- module_run(
  articles_files, 
  "stat_p_nonsig"
  )

exact_nonsig$table |> 
  count(text, sort = TRUE)
```

### `stat_effect_size` module

It detects reported standardized effect sizes for t-tests and f-tests

Number of t-tests and/or F-tests that did not report a standardized effect size
```{r}
es <- module_run(
  articles_files, 
  "stat_effect_size"
)

es$summary_text
```

Estimate number of F-tests not reporting an effect size
```{r}
head(es$table, 5)
es$table |> 
  mutate(as.factor(test)) |> 
  filter(test == "F-test") |> 
  count(is.na(es)) |> # returns cases where es == NA (not reported)
  mutate(prop = n / sum(n)) 
```
Estimate number of t-tests not reporting an effect size
```{r}
es$table |> 
  mutate(as.factor(test)) |> 
  filter(test == "t-test") |> 
  count(is.na(es)) |> # returns cases where es == NA (not reported)
  mutate(prop = n / sum(n)) 
```

Search for articles that have reported means ± SDs
```{r}
  search_text(
    articles_files,
    pattern = "(?i)mean\\s*±\\s*(sd|s\\.d\\.)", 
    section = "results",
    return = "sentence"
    )
```

Search for articles that might have reported 95% CI
```{r}
  search_text(
    articles_files,
    pattern = "(?i)95%\\s*(CI|confidence\\s*interval)",
    section = "results",
    return = "sentence"
    )
```

### `stat_check` module
```{r}
statcheck_p_values <- module_run(
  articles_files, 
  "stat_check"
  )

statcheck_p_values$table
```
`error` refers to p-values that have been misreported and `decision_error` refers to
when the error changes the decision

#### Prevalence of errors 
```{r}
statcheck_p_values$table |> 
  mutate(as.factor(error)) |> 
  count(error) |>         
  mutate(prevalence = round(n / nrow(all_ps) * 100, 1)) 
```

#### Prevalence of errors by decision errors
```{r}
statcheck_p_values$table |> 
  mutate(across(ends_with("error"), as.factor)) |> 
  count(error, decision_error) |>         
  group_by(error) |> 
  mutate(prevalence = round(n / sum(n) * 100, 1)) |> 
  ungroup()
```

#### Check for large discrepancies and possible false positives
```{r}
statcheck_p_values$table |> 
  mutate(flagged_p_values = round(computed_p / reported_p, 1)) |> 
           filter(flagged_p_values >= 2)
```

#### Check for large discrepancies and possible false positives only when `decision_error` == TRUE
```{r}
statcheck_p_values$table |> 
  filter(decision_error, round(computed_p / reported_p, 1) >= 2)
```
